---
title: "Assignment2_Part1_VoiceInSchizophrenia"
author: "Riccardo Fusaroli"
date: "July 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/home/tnoncs/Assignment3')
```

## Assignment 2 - Part 1 - Assessing voice in schizophrenia

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of Diagnosis. Can we do better with our new fancy complex skills?

The corpus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first Diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). I have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

N.B. Question to be answered via email to Celine: can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interpret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.

N.B. There are looots of files to be dealt with. Probably too many for your computer. This is a challenge for you. Some (complementary) possible strategies: You can select a subset of files only (and you have to justify your choice). You can learn how to use the apply() or map() functions. You can coordinate with classmates.

1. In the course of this assignment you have to first select one datafile and figure out how to:

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
```{r pressure}
getwd()
Studyone= read.table("Pitch/Study1D0S101T2_f0.txt", header=T)

demo_data= read.table("DemoData.txt", header=T)

mean=mean(Studyone$f0)
sd=sd(Studyone$f0)
range=range(Studyone$f0)
```
- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)

iqr like sd but doesn't assume the bell form
mean absolute deviation: the difference of a value from the mean - make it absolute value to remove minuses - get the mean of them
coef of variation= sd/mean
```{r pressure}
median=median(Studyone$f0)
iqr=IQR(Studyone$f0)
meanad= mean(abs(Studyone$f0 - mean(Studyone$f0))) # mean absolute deviation
# coefvar=sd/mean*100
coefvar=sd(Studyone$f0)/mean(Studyone$f0)*100
standard_data=cbind(mean, sd, range, median, iqr, meanad, coefvar) 
```

- Extract "complex" descriptors: recurrence quantification analysis

```{r pressure}
library(crqa)
#Finding the optimal parameters: 
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 
ans= optimizeParam(Studyone$f0, Studyone$f0, par, min.rec= 3.5, max.rec= 4.5)
#Choose a common value for delay, emb.dim and radius
Results=crqa (Studyone$f0, Studyone$f0, delay=ans$delay, embed=ans$emddim, radius=ans$radius, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)

# #Represent the plot: 
RP=Results$RP 
RP = matrix(as.numeric(RP), nrow = ncol(RP))  
cols = c("white","blue4") 
image(RP, xlab = "", ylab = "", col = cols) 

#Save the information from crqa
RR= Results$RR
DET= Results$DET
maxL=Results$maxL
L=Results$L
ENTR=Results$ENTR
LAM=Results$LAM
TT=Results$TT

rqa=data.frame(RR, DET, maxL, L, ENTR, LAM, TT)

all_data=merge(standard_data, rqa)
```

2. Second you will have to turn the code into a function and loop through all the files (or even better use apply/sapply/lapply)
- Remember to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)
```{r pressure}

file_list = list.files(path= "Pitch", full.names=T)
all_files=lapply(file_list, FUN=read.table, header=TRUE)

names(all_files)= substring(file_list, 7, 50) 

nombres=names(all_files)

#to merge
demo_data$Diagnosis=plyr::revalue(demo_data$Diagnosis, c("Control"="0", "Schizophrenia"="1"))
demo_data = plyr::rename(demo_data, c("Subject" = "id"))

#get out all basic stuff from all filenames
i= 1
basic_data=data.frame()

for (n in nombres) {
  Study= substring(nombres[i], 6, 6)
  Diagnosis= substring(nombres[i], 8, 8)
  id= substring(nombres[i], 10, 12)
  ido= unlist(regmatches(nombres[i], regexec("D(.+)T", nombres[i]))) 
  id_diff= ido[2]
  tr= unlist(regmatches(nombres[i], regexec("T(.+)_", nombres[i])))
  trial= tr[2]
  basic= cbind(Study, id, id_diff, trial, Diagnosis)
  basic_data=rbind(basic_data, basic) # save it, or it will replace the previous one 
  i=i+1
}

rm(basic, n, Study, Diagnosis, trial, id, i, tr)

basic_and_demo=merge(basic_data, demo_data)

```

#Code to deal with all the files at once

The goal is to have a file that contains basic- standard- rqa- and demodata, for all the voice files.
1. function to get basic data # from filenames. don't change the name of the folder they're in.
2. function to get standard data # basic calculations - mean, sd, etc...
3. function to have rqa data # ?
4. create list with all filenames
5. loop through the functions using the list
6. merge the result with demo data


```{r pressure}

#A function to get the info from the name of the files

get_name= function(f_name) {
  nombres= substring(f_name, 7, 50) #get rid of Pitch\
  
  Study= substring(nombres, 6, 6) # get the number after Study
  Diagnosis= substring(nombres, 8, 8) #get the number after D
  id= substring(nombres, 10, 12) #get the number after S #ids are the same for two person - wrong
  ido= unlist(regmatches(nombres, regexec("D(.+)T", nombres))) #create extra ids that has the Diagnosis too
  id_diff= ido[2] # "0S101" - the difference between the same ids is the Diagnosis, here number before S
  tr= unlist(regmatches(nombres, regexec("T(.+)_", nombres))) #extracts everything between T and _
  trial= tr[2] #get the number after T (using this method because there's trial 10 also)
  
  basic= data.frame(Study, id, id_diff, trial, Diagnosis) #save them to a dataframe
  return(basic)
}

```

```{r pressure}

#A function to get the standard data

get_standard= function(f_name) {
  nombres=read.table(f_name, header=T) 
  nev= nombres$f0                                       
  mean=mean(nev)
  sd=sd(nev)
  minimum= min(nev)
  maximum=max(nev)
  range= maximum - minimum
  median=median(nev)
  iqr=IQR(nev)
  meanad= mean(abs(nev - mean(nev)))
  coefvar=sd(nev)/mean(nev)*100
  standard=data.frame(mean, sd, minimum, maximum, range, median, iqr, meanad,coefvar) 
  return(standard)
}

```

```{r pressure}
# get the rqa data
library(crqa)

get_rqa= function(f_name) {
  nombres=read.table(f_name, header=T)
  
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100, radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0, whiteline = FALSE,  recpt = FALSE, fnnpercent = 10,  typeami = "mindip") 
  
  ans= optimizeParam(nombres$f0, nombres$f0, par, min.rec= 3.5, max.rec= 4.5)
  
  Results=crqa (nombres$f0, nombres$f0, delay=ans$delay, embed=ans$emddim, radius=ans$radius,
                normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
  RR= Results$RR
  DET= Results$DET
  maxL=Results$maxL
  L=Results$L
  ENTR=Results$ENTR
  LAM=Results$LAM
  TT=Results$TT
  
  rqa=data.frame(RR, DET, maxL, L, ENTR, LAM, TT)
  return(rqa)
}

#NAs if there's an error
empty=function(){
  RR= NA
  DET=NA
  maxL=NA
  L=NA
  ENTR=NA
  LAM=NA
  TT=NA
  empty= data.frame(RR, DET,maxL,L,ENTR,LAM,TT)
  return(empty)
}
  
please= try(get_rqa("Pitch/Study1D0S101T8_f0.txt"), silent = T)
if (class(please) == "try-error") { 
  msg <- geterrmessage()           #saves the error message
  please=empty() 
  }

```


```{r pressure}
# read in demo data
read_demo= function() {
  adat= read.table("DemoData.txt", header=T)
  adat$Diagnosis=plyr::revalue(adat$Diagnosis, c("Control"="0", "Schizophrenia"="1"))
  adat = plyr::rename(adat, c("Subject" = "id"))
  return(adat)
}

```


```{r pressure}
# DO NOT RUN IT!!!!!

file_list = list.files(path= "Pitch", full.names=T)

playa=data.frame() # empty dataframe to store data
for (i in file_list){
  basic=get_name(i) 
  stan= get_standard(i) 
  part= try(get_rqa(i), silent=T) 
  if (class(part) == "try-error") {
    msg <- geterrmessage()
    part=empty()
  }
  niidea= cbind(basic, stan, part) 
  playa= rbind(playa, niidea) #fill in the dataframe
}


playa= read_demo() # demo
port3_data= merge(niidea, playa) 

write.csv(port3_data, file = "port3_data.csv", row.names = F) 


```

```{r pressure}
# Data received FROM Viktória Torday

#read in data
port3_data= read.csv("port3_data.csv")
port3_data$diagnosis= as.factor(port3_data$diagnosis)

```


3. Make one model per acoustic feature and test whether you can observe significant difference due to Diagnosis. Tip: Which other fixed factors should you control for (that is, include in the model)? Which random ones?
- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 
```{r pressure}
#library(lme4)
library(lmerTest)

#using the model Riccardo created in class
#model_ric= range ~ Diagnosis + trial + (1 + trial | Study) + (1 + Diagnosis + trial | id)

model1= lmer(mean ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id) ,port3_data)
summary(model1) #significant

model2= lmer(sd ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model2) #not significant

model3= lmer(minimum ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model3) #significant 

model4= lmer(maximum ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model4) #not significant 

model5= lmer(median ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model5) #significant

model6= lmer(iqr ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model6) #not significant

model7= lmer(meanad ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model7) #not significant

model8= lmer(coefvar ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model8) #significant 
r.squaredGLMM(model8) 

model9= lmer(range ~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model9) # not significant 

model10= lmer(RR~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model10) # not significant

model11= lmer(DET~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model11) #significant

#High DET: when the system recurs, it tends to do that for at least two states in phase space
#Low DET: when the system recurs, it tends to do that for only 1 statE


model12= lmer(maxL~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model12) # not significant

model13= lmer(L~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model13) # not significant

model14= lmer(ENTR~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model14) #significant

#High Entropy: the system recurs in many different ways: sometimes short sequences, sometimes long sequences. 
#Low Entropy: the system recurs in stereotyped ways: always the same length

model15= lmer(LAM~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model15) #significant

#High LAM: sequences tend to be composed of the same state repeated over and over
#Low LAM: sequences if present (cf. DET) tend to be composed of different states

model16= lmer(TT~ diagnosis + trial + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model16) #significant

#High TT: when recurring a state the system tends to be trapped into it, repeating it again and again for long periods
#Low TT: the system tends not to repeat the same state
```

3a. Is study a significant predictor in these models? What should you infer from this? Does study interact with diagnosis? What should you infer from this?

```{r pressure}

model17= lmer(mean ~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id) ,port3_data)
summary(model17) # not significant

model18= lmer(mean ~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id) ,port3_data)
summary(model18) # not significant

model19= lmer(sd ~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model19) #not significant

model20= lmer(sd ~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model20) #not significant

model25= lmer(median ~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model25) # not significant

model26= lmer(median ~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model26)  #not significant

model27= lmer(iqr ~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model27) #not significant

model28= lmer(iqr ~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model28) #not significant

model29= lmer(meanad ~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model29) #not significant

model30= lmer(meanad ~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model30) #not significant

model31= lmer(coefvar ~ diagnosis + trial  + study +  (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model31) # not significant 

model32= lmer(coefvar ~ diagnosis + trial  + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model32) # not significant  

model33= lmer(range ~ diagnosis + trial  + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model33) # not significant 

model34= lmer(range ~ diagnosis + trial  + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model34) # significant ß= -13.0712  SE= 6.3840  p= 0.04341

model35= lmer(RR~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model35) # not significant

model36= lmer(RR~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model36) # not significant

model37= lmer(DET~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model37) # not significant

model38= lmer(DET~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model38) # not significant

model39= lmer(maxL~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model39) # significant  ß= 126.627  SE= 27.694  p< .0001 (1.97e-05)

model40= lmer(maxL~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model40) # not significant

model41= lmer(L~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model41) # significant ß= -0.25390 SE= 0.07765  p= 0.00142

model42= lmer(L~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model42) # not significant

model43= lmer(ENTR~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model43) # not significant

model44= lmer(ENTR~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model44) # not significant

model45= lmer(LAM~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model45) # not significant

model46= lmer(LAM~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model46) # not significant

model47= lmer(TT~ diagnosis + trial + study + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model47) # not significant

model48= lmer(TT~ diagnosis + trial + study*diagnosis + (1 + trial | study) + (1 + diagnosis + trial | id),port3_data)
summary(model48) # not significant

```

***

4. Bonus Question: Compare effect size of Diagnosis across the different measures. Which measure seems most sensitive?
- Tip: to compare across measures you need to put all of them on the same scale, that is, you need to "standardize" them (z-score)

5. Bonus question. In the Clinical Info file you have additional information about the participants. Which additional parameters (e.g. age, gender) should we control for? Report the effects.

6. Write a paragraph reporting methods and results

[Next assignment: can we use these measures to build a tool that diagnoses people from voice only?]

## N.B. Remember to save the acoustic features of voice in a separate file, so to be able to load them next time